---
title: "code-along-storefronts"
output: html_document
date: "2025-11-24"
---

Hi! Welcome to your first R Markdown. Here you can write text and add code to specific snippets, running them each individually.

Let's start by running our setup to get our libraries
```{r setup, include=FALSE}
library(tidyverse)
library(sf)

knitr::opts_chunk$set(echo = TRUE)
```

Let's come up with a research question!

I was wondering what areas of the city are doing well for commercial retail after the pandemic.

So I'm going to go look and see what data there is on business activity!

And noticed this dataset on [storefronts reported vacant](https://data.cityofnewyork.us/City-Government/Storefronts-Reported-Vacant-or-Not/92iy-9c3n/about_data) is on the city open data site

Let's grab that data for the past two years!

```{r}
# Define the API endpoint and query parameters
url <- "https://data.cityofnewyork.us/resource/92iy-9c3n.csv"
query <- paste0(url, "?$limit=1000000")

# Read the CSV file with the specified limit
storefronts <- read_csv(query)

```

This is a huge dataset! It would be really difficult to work with in QGIS. Let's use R to summarize some information we can map.

Let's start by cleaning the data a little bit.
- Get only the data from the last full calendar year.
- Reformat the borough_block_lot as a character column
```{r}

storefronts_clean <- storefronts #write your code here

```

Let's do some exploratory data analysis.

I wonder where all these properties are located? Should we try mapping it?

Here I'll help you out using the spatial features package to make the dataset into a spatial object

```{r}

storefronts_sf <- storefronts_clean %>% 
  #uh oh... did this not run? If this didn't run, look at the error, you might have missing values. Use this line to filter them out.
  st_as_sf(coords = c("longitude", "latitude"), remove = F)

```

Now let's try mapping it. We can do this with our old friend ggplot!

```{r}

storefronts_sf %>% 
  ggplot()+
  geom_sf(mapping = aes())

```
Looks like we have some zeroes for longitude and latitude. Filter them out and plot again.

```{r}

storefronts_sf %>% 
  #add a line here to filter
  ggplot()+
  geom_sf(mapping = aes())

```


Hey! looks like NYC. But that's a few too many points to look at.

Let's take a look at vacancy. Use count() to quickly see how many are vacant
```{r}
#write your code here

```


Now let's take that dataset and create a new object where we summarize the vacancy rate by community board on the last day of the year.

```{r}

storefronts_cb_sum <- storefronts_clean #write your code here

```

We could try writing out our data here, and joining it to shapes in QGIS. But what if we did that in R instead? We know how to do joins in R!

Let's use the sf package to read in a shapefile of the community districts(). We can go directly to sf by adding .geojson to the API endpoint.

```{r}

cds_sf <- read_sf("https://data.cityofnewyork.us/resource/ruf7-3wgc.geojson")

```

Or we can read it in as a csv and then set the geometry.

```{r}

cds_csv <- read_csv("https://data.cityofnewyork.us/resource/ruf7-3wgc.csv")

cds_sf <- cds_csv %>% 
  st_as_sf(wkt = "the_geom")#the name of the column where the geography info is!

```

Now let's join them together! Write a join to match our summary by cd to the shapes.

```{r}

cd_summary_sf <- #write your code here
  
#hint keep the shapefile on the left to preserve the spatial formatting of the object

```

Amazing! We can use sf to export this as a shapefile that has all the data we need, ready for QGIS

```{r}

write_sf(cd_summary_sf, "cd_summary.geojson")

```
But hey! We could also make this map in R! Let's try plotting our summarized shapefile with ggplot.

```{r}

cd_map <- cd_summary_sf %>% 
  st_set_crs("2263") %>% #I'm going to change the projection for you!
  #write your code here

cd_map

```

It could use a little styling... add a nicer color ramp and give it a theme

```{r}
cd_map_pretty <- cd_map +
  #write your code here

cd_map_pretty

```
I could even write this out to edit in a vector graphics software (you don't need to do this part in order to get credit!)

```{r}
ggsave("cd_summary_map.svg") #remember this always writes the last thing you ran

```


