---
title: "code-along-storefronts"
output: html_document
date: "2024-11-05"
---

Hi! Welcome to your first R Markdown. Here you can write text and add code to specific snippets, running them each individually.

Let's start by running our setup to get our libraries
```{r setup, include=FALSE}
library(tidyverse)
library(sf)

knitr::opts_chunk$set(echo = TRUE)
```

Let's come up with a research question!

I was wondering what areas of the city are doing well for commercial retail after the pandemic.

So I'm going to go look and see what data there is on business activity!

And noticed this dataset on [storefronts reported vacant](https://data.cityofnewyork.us/City-Government/Storefronts-Reported-Vacant-or-Not/92iy-9c3n/about_data) is on the city open data site

Let's grab that data for the past two years!

```{r}
# Define the API endpoint and query parameters
url <- "https://data.cityofnewyork.us/resource/92iy-9c3n.csv"
query <- paste0(url, "?$limit=1000000")

# Read the CSV file with the specified limit
storefronts <- read_csv(query)

```

This is a huge dataset! It would be really difficult to work with in QGIS. Let's use R to summarize some information we can map.

Let's start by cleaning the data a little bit.
- Get only the data from last year (2023). 2024 is not over yet
- Reformat the borough_block_lot as a character column
```{r}

storefronts_clean <- storefronts %>% 
  filter(reporting_year == "2023") %>% 
  mutate(borough_block_lot = as.character(borough_block_lot))

```

Let's do some exploratory data analysis.

I wonder where all these properties are located? Should we try mapping it?

Here I'll help you out using the spatial features package to make the dataset into a spatial object

```{r}

storefronts_sf <- storefronts_clean %>% 
  st_as_sf(coords = c("longitude", "latitude"), remove = F)

```

Now let's try mapping it. We can do this with our old friend ggplot!

```{r}

storefronts_sf %>% 
  ggplot()+
  geom_sf(mapping = aes())

```
Looks like we have some zeroes for longitude and latitude. Filter them out and plot again.

```{r}

storefronts_sf %>% 
  filter(latitude != 0) %>% 
  ggplot()+
  geom_sf(mapping = aes())

```


Hey! looks like NYC. But that's a few too many points to look at.

Let's take a look at vacancy. Use count() to quickly see how many are vacant
```{r}
#write your code here
storefronts_clean %>% 
  count(vacant_on_12_31)

storefronts_sf %>% 
  group_by(vacant_on_12_31) %>% 
  summarize(variable_name=n())

```


Now let's take that dataset and create a new object where we summarize the vacancy rate by community board on 12/31/2023.

```{r}

storefronts_cb_sum <- storefronts_clean %>% 
  group_by(community_board) %>% 
  summarize(count_per_board = n(),
            vacant_storefronts = sum(vacant_on_12_31 == "YES")) %>%
  mutate(vacancy_rate = vacant_storefronts/count_per_board*100,
         community_board = as.character(community_board))

storefronts_clean %>% 
  group_by(community_board, vacant_on_12_31) %>% 
  summarize(count_per_board = n()) %>%
  group_by(community_board) %>% 
  mutate(total = sum(count_per_board))

```

We could try writing out our data here, and joining it to shapes in QGIS. But what if we did that in R instead? We know how to do joins in R!

Let's use the sf package to read in a shapefile of the community districts(). We can go directly to sf by adding .geojson to the API endpoint.

```{r}

cds_sf <- read_sf("https://data.cityofnewyork.us/resource/jp9i-3b7y.geojson")

```

Or we can read it in as a csv and then set the geometry.

```{r}

cds_csv <- read_csv("https://data.cityofnewyork.us/resource/jp9i-3b7y.csv")

cds_sf <- cds_csv %>% 
  st_as_sf(wkt = "the_geom")#the name of the column where the geography info is!

```

Now let's join them together! Write a join to match our summary by cd to the shapes.

```{r}

cd_summary_sf <- left_join(cds_sf,storefronts_cb_sum, by = c("boro_cd"="community_board"))
  
#hint keep the shapefile on the left to preserve the spatial formatting of the object

```

Amazing! We can use sf to export this as a shapefile that has all the data we need, ready for QGIS

```{r}

write_sf(cd_summary_sf, "cd_summary.geojson")

```
But hey! We could also make this map in R! Let's try plotting our summarized shapefile with ggplot.

```{r}

cd_map <- cd_summary_sf %>% 
  st_set_crs("2263") %>%
  ggplot()+
  geom_sf(mapping = aes(fill = vacancy_rate), color = NA)
  #I'm going to change the projection for you!
  #write your code here

cd_map

```

It could use a little styling... add a nicer color ramp and give it a theme

```{r}
cd_map_pretty <- cd_map +
  scale_fill_gradient(low = "yellow", high = "red", na.value = "lightgray")+
  theme_void()

cd_map_pretty

```
I could even write this out to edit in a vector graphics software

```{r}
ggsave("cd_summary_map.svg") #remember this always writes the last thing you ran

```

